### 2014-08-29 paper/paper.Rmd
revised TVs section, added screenshot of shiny app, and revised play section

One of the factors in the PISA 2012 dataset is the number of televisions in the student's household. In the dataset, possible values were: "zero", "one", "two", or "three or more". Because there were so (relatively) few observations of "zero", for this invesitgation, the "zero" and "one" categories were combined to form a "one or less" category.

A linear regression as made for the students of each country, using the math score as the dependent variable and the number of televisions as the independent variable (as an ordered factor). Using R's `lm()` function, the direction and significance of the correlation was determined for each country. Significance is indicated by the p-value for a given regression being less than the popular 0.05.

@@

@@

```{r place_map_countries, echo=FALSE}
tags$div(
  style = "height: 400px",
  tags$p(
    tags$img(
      align = "right",
      src = "img/ijl_map.png",
      style="max-height: 100%"
    ),
    tags$img(
      align = "left",
      src = "img/ijl_countries.png",
      style="max-height: 100%"
    )
  )
)
```

Find your own correlations: http://bit.ly/pisa2012_explorer

@@

The PISA dataset contains dozens of household factors; we wanted a way to make our data-exploration more efficient. Moreover, we wanted a way that we could make explorations independent of each other, as during the course of the invesigation our team was in three cities over two hemispheres.

To make our exploration easier, we developed a "Shiny" app.

Shiny is a open-source tool, developed by RStudio, that allows an investigator to create a web-based appication powered by R.

>>>>

One of the factors in the data is the number of televisions in the student's household. Possible values were: "zero", "one", "two", or "three or more". Because there were so (relatively) few observations of "zero", for this invesitgation, the "zero" and "one" categories were combined to form a "one or less" category.

A linear model is fitted the students of each country, using the math score as the dependent variable and the number of televisions as the independent variable (as an ordered factor). The direction and significance of the correlation was determined for each country. Significance is indicated by the p-value for a given regression being less than the conventional 0.05.

@@

The figure below illustrates the relationship, along with another factor, student has a place to study. It is a complex diagram, but of interest is the middle column of plots - the boxplots overlaid on other information show the math scores of students relative to the number of TVs in the household, separately plotted for countries falling into one of three categories, negative, zero or  positive correlation. The white regions are violin plots of math scores by presence or absence of a place to study. A place to study universally leads to better math scores on average. The map is colored by correlation between number of TVs and math scores. The bar charts indicates sample size for the categories. The dot plots show average math scores by country, separately for each correlation group.

@@

!["TVs"](shiny-pisa.jpg)

@@

The PISA dataset contains hundreds of factors, and this article can describe just a few findings. The reader is welcome to explore the data themselves using an <a href="http://bit.ly/pisa2012_explorer"> interactive web tool</a>, made with <a href="http://www.rstudio.com">RStudio</a> and Shiny. Shiny is a open-source tool, developed by RStudio, that allows an investigator to create a web-based appication powered by R.
